Dynamics of Laterally Propagating Flames in X-ray Bursts. II. Realistic Burning \& Rotation

### 2021-02-26 xrb2/paper.tex
added variable definitions, minor grammar edits

\added{For the simple case of a laminar flame front driven by conduction, the propagation rate, 
$s_l$, of the flame is approximately
\begin{equation}
	\label{eqn:laminar_flame_speed}
	s_l \approx \left( \frac{\mathcal{D} \enucdot}{E} \right)^{\half}.
\end{equation}
Here, $\mathcal{D} = k_{th} / (\rho c_v)$ is the thermal diffusivity, $\enucdot$ is the specific
nuclear energy generation rate, and $E$ is the specific total energy \citep{timmeswoosley:1992}. 
\citet{cavecchi:2013} explored the effects of rotation on the flame's propagation, and noted that the 
flame speed in their simulations scaled with the ratio $L_R/H_0$ of Rossby length to atmospheric 
scale height. Combining this scaling with Equation \ref{eqn:laminar_flame_speed} yields a revised 
estimate for the flame speed, $s$:
\begin{equation}
	s \approx \frac{L_R}{H_0} s_l \approx \frac{L_R}{H_0} \left( \frac{\mathcal{D} \enucdot}{E} 
	\right)^{\half}.
\end{equation}
Observing that $L_R$ is inversely proportional to the angular rotation frequency $\Omega$ and 
and neglecting variations in $H_0$ and $E$, we can then derive the following approximate scaling 
relation:
\begin{equation}
	\label{eqn:flame_speed_scaling}
	s \appropto \frac{(\mathcal{D} \enucdot)^{\half}}{\Omega}.
\end{equation}
We were able to test the validity of this scaling relation in \citet{flame_wave1}, where we studied 
the effects of varying the thermal conductivity/diffusivity, energy generation rate, and rotation 
rate. These results supported the findings of \citet{cavecchi:2013}. Despite this formulation of 
the flame speed neglecting some relevant physics (e.g.\ turbulence), the relation proved
consistent with our simulation results.}

@@

\added{Rather than the fuel igniting in a 
localized hot spot and forming a flame which propagates across the domain, for this system the heat produced 
by the high temperature crust alone is sufficient to ignite material across the entire domain. Uniform burning 
across a large fraction of or even the entire surface of the neutron star has been observed in neutron stars 
with accretion rates that exceed the Eddington limit 
\citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}. However, it is important to note here that the
conditions in our model vary from the conditions that would occur in a neutron star ocean undergoing stable burning 
in a number of ways. Firstly, the model begins without any burning taking place, and initial model assumes that a 
significant amount of unburnt fuel has been able to build up. For a stably burning system, freshly accreted material 
would be continuously burnt, so there would be much less unburnt fuel available. Secondly, we are only modelling 
a few tens of milliseconds, rather than an ongoing process. We therefore do not demonstrate that this system would 
continue to burn stably if the high crustal temperature were maintained and sufficient fuel continued to be 
available, e.g.~through the continued accretion of fresh material. However, this run does suggest that 
a crust with a sufficiently high temperature can provide enough energy to heat the fuel above it to the point of 
ignition, and it may be that if these conditions were maintained that it would be able to sustain stable burning.}

@@

We ran a number of simulations of laterally propagating flames in XRBs in order 
to explore the effects of rotation and thermal structure. We found that increasing 
the rotation rate increased the energy generation rate within the flame and 
enhanced nuclear burning. Apart from the lowest rotation run (which failed to 
ignite), flame propagation was not noticeably impacted by rotation rate; by 
the time the different flame fronts reached steady propagation, they shared 
comparable velocities. These results are likely due to the rotation-dependent 
strength of the Coriolis force and its confinement of the flame balancing 
the enhanced nuclear burning. \added{This lack of dependence of the flame speed on 
the rotation rate could be tested using observations, as it infers that the burst 
rise time should have little dependence on the rotation rate. }

>>>>

\added{For the simple case of a laminar flame front driven by conduction, the propagation rate, 
$s_l$, of the flame is approximately
\begin{equation}
	\label{eqn:laminar_flame_speed}
	s_l \approx \left( \frac{\mathcal{D} \enucdot}{E} \right)^{\half}.
\end{equation}
Here, $\mathcal{D} = k_{th} / (\rho c_v)$ is the thermal diffusivity with thermal conductivity $k_{th}$ and specific heat capacity $c_v$, $\enucdot$ is the specific
nuclear energy generation rate, and $E$ is the specific total energy \citep{timmeswoosley:1992}. 
\citet{cavecchi:2013} explored the effects of rotation on the flame's propagation, and noted that the 
flame speed in their simulations scaled with the ratio $L_R/H_0$ of Rossby length to atmospheric 
scale height. Combining this scaling with Equation \ref{eqn:laminar_flame_speed} yields a revised 
estimate for the flame speed, $s$:
\begin{equation}
	s \approx \frac{L_R}{H_0} s_l \approx \frac{L_R}{H_0} \left( \frac{\mathcal{D} \enucdot}{E} 
	\right)^{\half}.
\end{equation}
Observing that $L_R$ is inversely proportional to the angular rotation frequency $\Omega$ and 
and neglecting variations in $H_0$ and $E$, we can then derive the following approximate scaling 
relation:
\begin{equation}
	\label{eqn:flame_speed_scaling}
	s \appropto \frac{(\mathcal{D} \enucdot)^{\half}}{\Omega}.
\end{equation}
We were able to test the validity of this scaling relation in \citet{flame_wave1}, where we studied 
the effects of varying the thermal conductivity/diffusivity, energy generation rate, and rotation 
rate. These results supported the findings of \citet{cavecchi:2013}. Despite this formulation of 
the flame speed neglecting some relevant physics (e.g.\ turbulence), the relation proved
consistent with our simulation results.}

@@

\added{Rather than the fuel igniting in a 
localized hot spot and forming a flame which propagates across the domain, for this system the heat produced 
by the high temperature crust alone is sufficient to ignite material across the entire domain. Uniform burning 
across a large fraction of or even the entire surface of the neutron star has been observed in neutron stars 
with accretion rates that exceed the Eddington limit 
\citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}. However, it is important to note here that the
conditions in our model vary from the conditions that would occur in a neutron star ocean undergoing stable burning 
in a number of ways. Firstly, the model begins without any burning taking place, and the initial model assumes that a 
significant amount of unburnt fuel has been able to build up. For a stably burning system, freshly accreted material 
would be continuously burnt, so there would be much less unburnt fuel available. Secondly, we are only modelling 
a few tens of milliseconds, rather than an ongoing process. We therefore do not demonstrate that this system would 
continue to burn stably if the high crustal temperature were maintained and sufficient fuel continued to be 
available, e.g.~through the continued accretion of fresh material. However, this run does suggest that 
a crust with a sufficiently high temperature can provide enough energy to heat the fuel above it to the point of 
ignition, and it may be that if these conditions were maintained that it would be able to sustain stable burning.}

@@

We ran a number of simulations of laterally propagating flames in XRBs in order 
to explore the effects of rotation and thermal structure. We found that increasing 
the rotation rate increased the energy generation rate within the flame and 
enhanced nuclear burning. Apart from the lowest rotation run (which failed to 
ignite), flame propagation was not noticeably impacted by rotation rate; by 
the time the different flame fronts reached steady propagation, they shared 
comparable velocities. These results are likely due to the rotation-dependent 
strength of the Coriolis force and its confinement of the flame balancing 
the enhanced nuclear burning. \added{This lack of dependence of the flame speed on 
the rotation rate could be tested using observations, as it implies that the burst 
rise time should have little dependence on the rotation rate. }
